---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# New column-pair mapping functions

```{r include=FALSE}
# Dev only: Load pairmaps from within pairmaps
devtools::load_all(".")
```

<!-- badges: start -->

[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental) [![CRAN status](https://www.r-pkg.org/badges/version/pairmaps)](https://CRAN.R-project.org/package=pairmaps)

<!-- badges: end -->

The pairmaps package provides `as_colpair_mapper()`, which modifies a function so that it is applied to each pair of columns in a data frame. Every function created this way is a wrapper around `corrr::colpair_map()`.

You can avoid pairmaps as a dependency by copying and pasting the functions returned by `as_colpair_mapper()` into your source code.

pairmaps is an unofficial add-on to [corrr](https://corrr.tidymodels.org/). It doesn't have any dependencies beyond those of corrr (and corrr itself).

## Installation

You can install the development version of pairmaps from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("lhdjung/pairmaps")
```

## Get started

Go to `vignette("using-pairmaps")`.

## Acknowledgements

pairmaps was inspired by [factory](https://github.com/jonthegeek/factory).
