---
title: "Adverb, function operator, or function factory?"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Adverb, function operator, or function factory?}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r include=FALSE}
# Dev only: Load pairmaps from within pairmaps
devtools::load_all(".")
```

Functions like `as_colpair_mapper()` are sometimes called *adverbs* because they modify the behavior of existing functions (i.e., "verbs"). Other examples are `purrr::safely()` and [`widyr::widely()`](https://juliasilge.github.io/widyr/reference/widely.html).

A more technical term is *function operator*, a special kind of *function factory*. The latter refers to any function that returns a function. In this way, function operators are simply function factories that take one or more functions as input (Wickham 2019, [ch. 10-11](https://adv-r.hadley.nz/function-factories.html)).

### Balancing pros and cons

I tend to prefer "function factory" because it's more formal than "adverb" but also highly evocative. It's used in the context of other languages as well, such as Python and Javascript. The term is less precise than "function operator", but I think this won't do much harm. Although function operators may not have the same use cases as other factories, the differences are not too significant during development. This is also true when working with the [factory package](https://github.com/jonthegeek/factory), a great tool for writing new function factories.

Yet "adverb" is an elegant metaphor that smoothly conveys these functions' usage. Consider this example from purrr:

```{r}
safe_log <- purrr::safely(log)
safe_log(10)
```

Somebody who is familiar with `log()` but not with function factories will more likely make sense of the idea that `safely()` modifies the way `log()` works. It might be less intuitive that `safely()` takes `log` as an input, returns the new `safe_log` function,[^1] and then calls that function with some other object as an argument --- although that is exactly what happens, and what function operators in general do.

[^1]: I leave out the parentheses here to emphasize that `log` and `safe_log` are functions, not function calls; and that they can be passed and returned like any other R object. Everywhere else, however, I use the conventional `log()`-like notation.

Accessible explanations are needed even more acutely if we rewrite the example:

```{r}
purrr::safely(log)(10)
```

This syntax --- two sets of parentheses, seemingly following the same function --- will puzzle anybody who encounters it for the first time. What goes on here is easier to explain by saying that `safely()` is an adverb that changes they way `log()` works. Note that the output is the same as before, so the two notations are equivalent.

Referring to such functions as "adverbs", however, is not very precise. R objects usually can't be modified in place (Wickham 2019, [ch. 2.3](https://adv-r.hadley.nz/names-values.html#copy-on-modify)), and adverbs don't actually change their input functions. Instead, they create and return entirely new functions that apply the input ones in some particular way. The sheer complexity of all this means that speaking of adverbs and modification will often be reasonable, especially when the focus is on *using* function factories, not writing them.

### Conclusion

All things considered, it seems wise of the developer community to speak of adverbs in user-facing contexts and of function factories / operators in developer-facing contexts. Both sources linked above that reference adverbs are documentation pages for existing functions. By contrast, chapters 10-11 of *Advanced R* (Wickham 2019) are directed at more experienced R programmers who might want to write function factories themselves. The same applies to the [factory package's readme](https://github.com/jonthegeek/factory). This audience will need a deeper understanding of the way such functions work, and the concept of "function factories" will bring them into sharper view.

*Advanced R* only uses the term "adverb" once, describing function operators in the [introduction](https://adv-r.hadley.nz/fp.html) to the relevant chapters: "They are like adverbs, because they typically modify the operation of a function." Isn't that a nice segue?

Sticking to the same few terms is advisable in any case, lest somebody points out that `as_colpair_mapper()` is actually an adverbial phrase.

### Citation dummy {style="display: none;"}

*(Here is an extra section. Its display is suppressed via the CSS element `display: none;` because its only purpose is to generate the citation @wickham2019 --- I want to style the citation in the text above more flexibly than the format allows (including a URL) but still get the reference pasted at the bottom.)*

## References
