---
title: "Using pairmaps"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using pairmaps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r include=FALSE}
# Dev only: Load pairmaps from within pairmaps
devtools::load_all(".")
```

This vignette shows you when and how to use pairmaps.

```{r setup, eval=FALSE}
library(pairmaps)
```

### Basics

The main function of pairmaps is the adverb `as_colpair_mapper()`. It modifies a given function --- or "verb" --- so that it operates on each pair of columns in a data frame.

More precisely, `as_colpair_mapper()` creates wrapper functions around `corrr::colpair_map()`. Each wrapper is bound to one specific other function and applies it to a data frame. For example:

```{r}
max_sum <- function(x, y) {
  max(sum(x), sum(y))
}
max_sum_map <- as_colpair_mapper(f = max_sum)
max_sum_map(mtcars)
```

The mapped function, `f`, must take at least two arguments (or the dots, `...`). The first two will be the column pairs, so the function must accept vectors for them.

Use `as_colpair_mapper()` if you plan to reuse the resulting function, like `max_sum_map()`, multiple times. A new wrapper function would add little value if you only want to use it once. In this case, directly calling `corrr::colpair_map()` might be the better option. Read more about `colpair_map()` in [this blogpost](https://www.tidyverse.org/blog/2020/12/corrr-0-4-3/#colpair_map).

`vary()` and `covary()` are built-in examples for useful functions made by `as_colpair_mapper()`.

### More with corrr

Running your data through `colpair_map()` --- directly or via a wrapper function made by `as_colpair_mapper()` --- will always return a `cor_df` tibble. `cor_df` is the same format as a data frame of correlations returned by `corrr::correlate()`, but it's not limited to computing correlations. Instead, `colpair_map()` arranges its results in this special format after applying any user-specified function which takes two or more arguments.

This enables you to plug your analysis into any R workflow focused on data frames, but especially the corrr package's arsenal of exploratory functions. Read about them in the [*Using corrr* vignette](https://corrr.tidymodels.org/articles/using-corrr.html).

### Defaults for defaults

You can control which defaults the `.diagonal` and `.quiet` arguments of the output function will have by specifying the `default_diagonal` and `default_quiet` arguments of `as_colpair_mapper()`:

```{r}
max_sum_map <- as_colpair_mapper(
  f = max_sum, default_diagonal = 1, default_quiet = TRUE
)

# The diagonal is `1`, and no message is displayed!
max_sum_map(mtcars)
```

The defaults *for these defaults* within `as_colpair_mapper()`, in turn, correspond to this:

```{r}
max_sum_map <- as_colpair_mapper(
  f = max_sum, default_diagonal = NA, default_quiet = FALSE
)
```

### Assignment and copying

#### Not evaluating too early

`as_colpair_mapper()` has one more argument, the Boolean `eval_f`. If you set it to `FALSE`, the input function will appear within the output function by name only. Look at `.f = max_sum` here:

```{r}
as_colpair_mapper(f = max_sum, eval_f = FALSE)
```

This is useful if you want to copy and paste the printed output into a script, especially within a package. Always set `eval_f` to `FALSE` if you *call* `as_colpair_mapper()` inside a package. (See below for non-package use cases.) For example, `vary()` and `covary()` are defined like this within pairmaps:

```{r eval=FALSE}
vary <- as_colpair_mapper(stats::var, eval_f = FALSE)
covary <- as_colpair_mapper(stats::cov, eval_f = FALSE)
```

The default, `eval_f = TRUE`, would replace the name of `max_sum()` by its function definition --- i.e., its list of arguments along with the entire function body:

```{r}
as_colpair_mapper(f = max_sum)
```

You can imagine that this might be confusing with longer functions than `max_sum()`. However, there is a deeper problem still: The factory-made function would only reflect the version of `max_sum()` at the specific point in time when you run `as_colpair_mapper()`. If `max_sum()` changes later on, you would then have to manually update the copied function, even when working within a package. Otherwise, it will rely on an outdated version of `max_sum()`!

If you aren't developing a package but copying the output of `as_colpair_mapper()` into some other script, you will have to manually update it anyways if you run `as_colpair_mapper()` with `eval_f = TRUE`.

#### Not evaluating too late

The purpose of the default `eval_f = TRUE` is to prevent the reverse problem. If you don't copy and paste the output function but assign it to a variable straight away, setting `eval_f` to `FALSE` might lead to a subtle issue with lazy evaluation (Wickham 2019, [ch. 10.2.3](https://adv-r.hadley.nz/function-factories.html#forcing-evaluation)):

```{r}
max_sum <- function(x, y) {
  max(sum(x), sum(y))
}
max_sum_map <- as_colpair_mapper(f = max_sum, eval_f = FALSE)
max_sum <- function(...) {
  sum1 <- sum(iris$Petal.Width)
  sum2 <- sum(mtcars$hp)
  sum3 <- sum(sunspots)
  max(sum1, sum2, sum3)
}
max_sum_map(mtcars)
```

What happened here? The `f = max_sum` argument was only evaluated when the output function, `max_sum_map()`, was ultimately called. Before it was, however, the value of `max_sum()` changed to a different function, which was then applied in place of the original `max_sum()`.

I made the point very apparent because the new `max_sum()` function returns a constant output when the tibble should include many different values. In more realistic cases, the problem might be harder to spot, and it would be safer to evaluate `f` to the original value of `max_sum()` right when `as_colpair_mapper()` creates `max_sum_map()`. The default `eval_f = TRUE` provides this safety.

If the new value can't be interpreted as a function, the bug doesn't occur but an error is thrown:

```{r, error=TRUE}
max_sum <- function(x, y) {
  max(sum(x), sum(y))
}
max_sum_map <- as_colpair_mapper(f = max_sum, eval_f = FALSE)

# This new `max_sum` is numeric, not a function:
max_sum <- max(sum(iris$Petal.Length, iris$Petal.Width))
max_sum_map(mtcars)
```

### Retrieving the mapped function

In case you have a pairmaps data frame but you aren't sure which function was applied to the column pairs, you can call `get_mapped_function()` with that data frame. Note that it will return the function as it is defined the moment you call `get_mapped_function()`. This means you can't use it to retrieve any earlier functions by the same name.

```{r}
max_sum <- function(x, y) {
  max(sum(x), sum(y))
}
max_sum_map <- as_colpair_mapper(f = max_sum)
df <- max_sum_map(mtcars)

get_mapped_function(df)
```

If you need the name of the mapped function, call `get_mapped_function_name()` instead:

```{r}
get_mapped_function_name(df)
```

Sometimes, you may not be sure if a data frame was returned by a function which, in turn, was made by `as_colpair_mapper()`. Test this with `is_pairmaps_df()`:

```{r}
is_pairmaps_df(df)
is_pairmaps_df(iris)
```

`get_mapped_function()` and `get_mapped_function_name()` internally call `is_pairmaps_df()` to make sure they actually deal with such a data frame. If the test fails, they will throw an error.

### Citation dummy {style="display: none;"}

*(Here is an extra section. Its display is suppressed via the CSS element `display: none;` because its only purpose is to generate the citation @wickham2019 --- I want to style the citation in the text above more flexibly than the format allows (including a URL) but still get the reference pasted at the bottom.)*

## References
